### 2018-03-24 tex/paper.tex
initial work on commentary; more to do

@@

In doing this we are required to refine the model of contracts in a number of ways in order to fit with a radically different context. 
\begin{itemize}
\item In the traditional model, enforcement of the contract is the responsibility of the legal system; in the decentralised blockchain model, in which there is no central authority, the contract needs to be enforced by design. 
\item As noted by  the authors of Findel~\cite{findel} this means in particular that payments need to be bounded; this is made concrete in our account by our notion of commitments.
\item There is not a single blockchain model, and we give a brief overview of the different mechanisms for scripting on blockchain (although this is not the primary purpose of this article). 
\begin{itemize}
\item
A particularly important technical distinction here is that between an account-style model as used by Ethereum~\cite{EthereumRationale}, and the unspent transaction output (UTxO) model of bitcoin~\cite{sok}. 
\item There is the question of how precisely the execution of the contract interacts with the blockchain, and how information is to be recorded on the chain.
\item Does the execution ``push'' for changes  to happen, or do external actors have to ``pull'' to effect changes enabled by execution of the contract?
\end{itemize}
\item \sjtnote{What else do we want to say here?}
\end{itemize}

@@

\sjtnote{To talk about, in no particular order \ldots}
\begin{itemize}
\item
Marlowe as the ``canary in the coalmine'': test out how it will work, how it will interface with
other aspects of the Cardano platform
\item
Be clear that this paper describes version 1.2 of Marlowe.
\end{itemize}

@@

A common feature of many blockchain platforms is that they provide multiple scripting languages. As we noted earlier, Ethereum can be programmed at the EVM level as well as by using the high-level Solidity language. Tezos~\cite{tezos-white-paper} supports the stack-based, strongly-typed, functional language Michelson, but also provides a high-level language, Liquidity, that compiles into Michelson. Liquidity is also functional and strongly typed, but provides more constructs in a more familiar syntax, namely that of a subset of OCaml. The \ae{}ternity system~\cite{aeternity} provides multiple languages and VMs:\footnote{As presented by Erik Stenman at CODE BEAM SF, March 2018, \url{https://www.youtube.com/watch?v=VXsqvfPIdWg}} the functional language  Sophia (akin to Reason) and the Functional Typed Warded Virtual Machine (FTWVM) for safe ``system level programming'', the language Varna and the HLM for simple contracts, and a (port of) Solidity and the EVM for compatibility with Ethereum. In a similar way, Cardano provides support for IELE~\sjtnote{citation}, a rational reconstruction of the EVM, and thus for Solidity contracts too.

@@

There are a variety of other approaches to scripting, including business process notations, such as the OMG's BPMN; process calculi such as Rholang; logic rules (in Formal Contract Logic), and finite state machines. More details of these are provided in the survey~\cite{cryptoeprint:2016:1156}.

@@

Some chains have a layered design, and in particular Cardano~\cite{cardano} has. The settlement layer (SL) is to support settlement of transactions but nothing more complicated, whereas the computation layer (or CL) supports more powerful general computations over the chain. Moreover, the Cardano roadmap \cite{cardano-rationale} envisages the possibility of computation taking place, in part at least, on a sidechain rather than on the main chain itself, which may support only the SL. Other off-chain approaches include the Lightning network, overlaid on blockchains (usually Bitcoin) and the channels of \ae{}ternity, introduced below.

@@

\sjtnote{To be completed.}

\section{The model}
The DSL is modelled as an algebraic type in Haskell, together with an executable small-step semantics. 
A running contract will interact with its environment in two ways.

@@

@@

@@

@@

The model is based on a number of assumptions.

\begin{itemize}
\item It is assumed that cryptographic ``services'' or operations are provided by a layer external to the system, and so they need not be modelled explicitly here.
\item Making a commitment is not something that a contract can perform; rather, it can require that a commitment is made, but that then has to be established externally: hence the input of (a set of) commitments at each step. The same observation also applies to redeeming a commitment of cash.
\item Observables are observed at a particular time and in a particular context. We assume that the recording and 
identification of each (instance of an) observation is provided by the infrastructure of the system. We provide a 
``little language'' for describing observations: \haskellinline{Observation}. Similar properties are true for
\haskellinline{Money} amounts: their values are calculated depending on the inputs and state of execution of the 
contract, they can be deterministically established given the context, and they are described through a small 
``expression language''.
\item The fact of making commitments also needs to be recorded.
\item It is important to be aware that there may be different instances of contracts executed, and for each instance there will typically be different instances of observables (and commitments) observed and recorded.
\item The model manages the release of funds back to the committer when a cash commitment expires (see discussion of the \haskellinline{fullStep} function below).
\item We assume that time is ``coarse grained'' and measured simply by block number, so that timeouts etc.\ are delimited using nothing finer than block number. 
\end{itemize}

\medskip
\noindent
The type of contracts is given by

@@

@@

The constructors should be interpreted thus:

@@

\item \haskellinline{When obs timeout k def_k} This contract will not progress until \haskellinline{obs} is \haskellinline{True} or until the current block number is greater than or equal to the one specified by timeout. In case the timeout applies, the contract will continue as \haskellinline{def_k}, if the timeout does not apply and \haskellinline{obs} is \haskellinline{True}, then the contract continues as \haskellinline{k}. Otherwise the contract is quiescent.
\end{itemize}

@@

@@

We should monadify, so that identifiers are automatically generated. This gives two nice properties. First, they are guaranteed to be unique; secondly they can be generated only in cases where a commit has been made successfully (within the required time period). It might well be that these should be generated by the system; this could guarantee uniqueness, freshness etc, and could also lead to a monadic DSL, with bind used for binding names to commitments.

\section{Analysis}
What questions can we ask? What answers can we give?

\subsubsection{What are the valid contracts?}

One possible definition is ``those contracts that do not lead to a \haskellinline{FailedPay} action''. Does this make sense independently of the inputs to a contract (i.e. observables, commitments, redeems, timeouts, and randomness)?
Some relevant points (NB discussion incomplete):
\begin{itemize}
\item For \haskellinline{Pay} require a sufficient balance not to produce a \haskellinline{FailedPay} or a \haskellinline{DuplicateRedeem} action. 
\item For \haskellinline{Both} require a sufficient balance for both contracts combined. Need to think here whether the two conjuncts should be independent, or whether a commitment in one half can be used in the other.
\item For \haskellinline{Choice} require a sufficient balance for the maximum of the requirements of the individual contracts. 
\item For \haskellinline{When}, need to ``age'' the existing commitments relative to the new starting point of the continuation contract. 
\item Are there any constraints on  \haskellinline{CommitCash} and \haskellinline{RedeemCC} statements? Not directly, but there need to be sufficient funds available to pay if needed, or enough time for the commitments to time out, if not.
\item Originally we had included that we should include all the ``progress'' requirements from point 2 below; but that's probably not the right thing here.

\end{itemize}

\subsubsection{Do all contracts potentially evaluate to \haskellinline{Null}?}

That is, for each contract do there exist sequences of values of Commit and OS such that stepping the contract with these inputs leads to Null? Might we have to rule out unsatisfiable observations?
Some relevant points (NB discussion incomplete):
\begin{itemize}
\item For \haskellinline{When}, need to be able to make \haskellinline{interpretObs st obs os} true. Since there is a timeout to \haskellinline{When} then there's no need to impose this condition, as we'll progress at timeout if not before.
\item For \haskellinline{CommitCash}, need to have the right commits in the \haskellinline{Commit} argument.
\item Similarly for \haskellinline{RedeemCC} statements.
\item It would also be possible to put timeouts on to \haskellinline{CommitCash},  and \haskellinline{RedeemCC} operations, and if this were to be done there would be no remaining requirements for contract execution to make progress.
\item There are no constraints on other constructs.
\end{itemize}

\section{Questions}

@@

\section{Related work}

\begin{itemize}
\item Original SPJ paper~\cite{PeytonJones:2000}, and multi-party successor. Original Szabo paper?
\item Findel~\cite{findel}
\end{itemize}

In this section, we explain how the DSL can be used for designing a contract that incentivises saving. We use the 
following example: ``participant 2'' offers 20 ADA to ``participant 1'' in exchange for saving 100 ADA for 100 blocks. 
``Participant 1'' will be able to withdraw the savings after block 20, but if that option is chosen, then ``participant 
2'' will not pay the extra 20 ADA to ``participant 1''.

The whole contract remains as follows:
\begin{haskellcode}
CommitCash (IdentCC 1) 1
           (ConstMoney 100)
           10 200
           (CommitCash (IdentCC 2) 2
                       (ConstMoney 20)
                       20 200
                       (When (PersonChoseSomething (IdentChoice 1) 1)
                             100
                             (Both (RedeemCC (IdentCC 1) Null)
                                   (RedeemCC (IdentCC 2) Null))
                             (Pay (IdentPay 1) 2 1
                                  (ConstMoney 20)
                                  200
                                  (Both (RedeemCC (IdentCC 1) Null)
                                        (RedeemCC (IdentCC 2) Null))))
                       (RedeemCC (IdentCC 1) Null))
           Null
\end{haskellcode}

\subsection{Response to comments on the original document made by Charles on Slack}

>>>>

\setlength{\itemsep}{2pt}

@@

In doing this we are required to refine the model of contracts in a number of ways in order to fit with a radically different context. 
\begin{itemize}
\item In the traditional model, enforcement of the contract is the responsibility of the legal system; in the decentralised blockchain model, in which there is no central authority, the contract needs to be enforced by design. 
\item There is not a single blockchain model, and we give a brief overview of the different mechanisms for scripting on blockchain (although this is not the primary purpose of this article). 
\begin{itemize}
\item
A particularly important technical distinction here is that between an account-style model as used by Ethereum~\cite{EthereumRationale}, and the unspent transaction output (UTxO) model of bitcoin~\cite{sok}. 
\item There is the question of how precisely the execution of the contract interacts with the blockchain, and how information is to be recorded on the chain.
\item Does the execution ``push'' for changes  to happen, or do external actors have to ``pull'' to effect changes enabled by execution of the contract?
\end{itemize}
\item \sjtnote{What else do we want to say here?}
\end{itemize}

@@

\sjtnote{To talk about, in no particular order \ldots}
\begin{itemize}
\item
Marlowe as the ``canary in the coalmine'': test out how it will work, how it will interface with
other aspects of the Cardano platform
\item
Be clear that this paper describes version 1.2 of Marlowe.
\item 
Enforcement and promptness.
\end{itemize}

@@

A common feature of many blockchain platforms is that they provide multiple scripting languages. As we noted earlier, Ethereum can be programmed at the EVM level as well as by using the high-level Solidity language. Tezos~\cite{tezos-white-paper} supports the stack-based, strongly-typed, functional language Michelson, but also provides a high-level language, Liquidity, that compiles into Michelson. Liquidity is also functional and strongly typed, but provides more constructs in a more familiar syntax, namely that of a subset of OCaml.

The \ae{}ternity system~\cite{aeternity} provides multiple languages and VMs:\footnote{As presented by Erik Stenman at CODE BEAM SF, March 2018, \url{https://www.youtube.com/watch?v=VXsqvfPIdWg}} the functional language  Sophia (akin to Reason) and the Functional Typed Warded Virtual Machine (FTWVM) for safe ``system level programming'', the language Varna and the HLM for simple contracts, and a (port of) Solidity and the EVM for compatibility with Ethereum. In a similar way, Cardano provides support for IELE~\sjtnote{citation}, a rational reconstruction of the EVM, and thus for Solidity contracts too.

@@

There are a variety of other approaches to scripting, including business process notations, such as the OMG's BPMN; process calculi such as Rholang; logic rules (in Formal Contract Logic), and finite state machines. More details of these are provided in~\cite{cryptoeprint:2016:1156}.

\paragraph{Related work}

The Findel~\cite{findel} project examines financial contracts on the Ethereum platform, and the authors note that  payments need to be bounded; this is made concrete in our account by our notion of commitments. They take no account of commitments or timeouts as we do below, but it should be noted that the Ethereum platform is more powerful than the one that we target in this paper.

@@

Some chains have a layered design, in particular Cardano~\cite{cardano}. The Cardano Settlement Layer (SL) is to support settlement of transactions but nothing more complicated, whereas the Computation Layer (or CL) supports more powerful general computations over the chain. Moreover, the Cardano roadmap \cite{cardano-rationale} envisages the possibility of computation taking place, in part at least, on a sidechain rather than on the main chain itself, which may support only the SL. Other off-chain approaches include the Lightning network, overlaid on blockchains (usually Bitcoin) and the channels of \ae{}ternity, introduced below.

@@

In designing Marlowe we have chosen a number of options; this description applies to Marlowe version 1.2 \sjtnote{Is there an obvious way of making that clear in the repo? I really don't understand how git deals with ``releases'' or versions.}

\paragraph{Pull}

While earlier versions used a ``push'' model of operations, we have revised the language to give a ``pull'' semantics. This means, for example, that the language can \emph{enable} a payment to be made, but cannot make the payment happen. It is up to the interested party (or parties) to make the change happen. It is therefore possible for the payment not to happen, despite the fact that this is against the interest of the receiving party. The same

\paragraph{UTxO vs.\ accounts} \sjtnote{Pablo: please check this section carefully}

The language semantics is described in terms of \emph{commitments} of sums of cryptocurrency -- which we choose to abbreviate to ``cash'' -- and so is agnostic of the underlying implementation.  However, the current redemption mechanism in Plutus Core is not strong enough to support Marlowe, and we are actively investigating the best way to extend this mechanism to implement Marlowe in full on Plutus Core.

\paragraph{DSL or EDSL?}

Marlowe is described as an algebraic \haskellinline{data} type in Haskell, and as such can be seen as a simple EDSL. Note, however, that the implementation presented here is for the purposes of presenting the semantics and behaviour of the language in the abstract. The implementation on the Cardano platform could be stand-alone, but more likely it will be a DSL embedded in Plutus or Plutus Core. The details of this remain to be decided, as noted in the introduction.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.6\textwidth]{pix/context.png}
\caption{The context for the contract}
\label{fig:context}
\end{center}
\end{figure}

\section{The model}

The DSL is modelled as an algebraic type in Haskell, together with an executable small-step semantics. 
We start by looking at the different types used by the model, and then look at the \haskellinline{Contract} DSL itself. \sjtnote{To complete this introduction}

\subsection{The model types}

A running contract will interact with its environment in two ways, as shown in Figure \ref{fig:context}.

\paragraph{Observables}

@@

\paragraph{Inputs}

@@

\paragraph{Commitments}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\textwidth]{pix/step-type.png}
\caption{The \haskellinline{step} function}
\label{fig:step-function}
\end{center}
\end{figure}

\paragraph{The \haskellinline{step} function}

@@

which is also shown in Figure \ref{fig:step-function}.

\paragraph{Infrastructure}

The model makes a number of assumptions about the blockchain infrastructure in which it is run.

\begin{itemize}
\item It is assumed that cryptographic functions and operations are provided by a layer external to the system, and so they need not be modelled explicitly here.
\item We assume that time is ``coarse grained'' and measured simply by block number, so that timeouts etc.\ are delimited using nothing finer than block number. 
\item Because we assume a ``pull'' model, making a commitment is not something that a contract can perform; rather, it can enable that a commitment is made, but that then has to be established externally: hence the input of (a set of) commitments at each step. The same observation also applies to redeeming a commitment of cash. Commitments that have been made or redeemed need to be recorded somehow.

\item Observables are observed at a particular time and in a particular context. We assume that the recording and 
identification of each (instance of an) observation is provided by the infrastructure of the system. We provide a 
``little language'' for describing observations: \haskellinline{Observation}. 
\item
Similar properties are true for
\haskellinline{Money} amounts: their values are calculated depending on the inputs and state of execution of the 
contract, they can be deterministically established given the context, and they are described through a small 
``expression language''.
\item It is important to be aware that there may be different instances of contracts executed, and for each instance there will typically be different instances of observables (and commitments) observed and recorded.
\item The model manages the release of funds back to the committer when a cash commitment expires (see discussion of the \haskellinline{fullStep} function below).
\end{itemize}

\subsection{The \haskellinline{Contract} type}

The type of contracts is given by

@@

\subsection{The \haskellinline{step} function}

@@

The constructors are interpreted like this:

@@

\item \haskellinline{When obs timeout k def_k} This contract will not progress until \haskellinline{obs} is \haskellinline{True} or until the current block number is greater than or equal to the one specified by timeout. In case the timeout applies, the contract will continue as \haskellinline{def_k}, if the timeout does not apply and \haskellinline{obs} is \haskellinline{True}, then the contract continues as \haskellinline{k}. Otherwise the contract is quiescent.
\end{itemize}
We look next at an example of Marlowe in action.

In this section, we explain how Marlowe can be used for designing a contract that incentivises saving. We use the 
following example: ``participant 2'' offers 20 ADA to ``participant 1'' in exchange for saving 100 ADA for 100 blocks. 
``Participant 1'' will be able to withdraw the savings after block 20, but if that option is chosen, then ``participant 
2'' will not pay the extra 20 ADA to ``participant 1''.

The whole contract remains as follows:
\begin{haskellcode}
CommitCash (IdentCC 1) 1
           (ConstMoney 100)
           10 200
           (CommitCash (IdentCC 2) 2
                       (ConstMoney 20)
                       20 200
                       (When (PersonChoseSomething (IdentChoice 1) 1)
                             100
                             (Both (RedeemCC (IdentCC 1) Null)
                                   (RedeemCC (IdentCC 2) Null))
                             (Pay (IdentPay 1) 2 1
                                  (ConstMoney 20)
                                  200
                                  (Both (RedeemCC (IdentCC 1) Null)
                                        (RedeemCC (IdentCC 2) Null))))
                       (RedeemCC (IdentCC 1) Null))
           Null
\end{haskellcode}
In the next section we look again at the model and the Marlowe language.

\section{Reflection}
We have given a concrete model and DSL, namely Marlowe version 1.2. In this section we examine how the model can be taken further, and reflect on different possible choices that might be made.

@@

\paragraph{Value commitments}

@@

\paragraph{A monadic DSL}

Haskell is a pure functional programming language, and uses the \haskellinline{Monad} type class to encapsulate operations that can potentially have a side effect. Many introductions to monads exist \sjtnote{cite some} but for the sake of this discussion it is enough to think of a value of type \haskellinline{m a} where \haskellinline{m} is a monad as a ``computation that results in something of type \haskellinline{a}''. The \haskellinline{Monad} API allows such computations to be assembled, and then once assembled they may be run, in general causing some side-effects as a part of the computation.

We can make the \haskellinline{Contract} type monadic, so that new, unique identifiers are automatically generated. This gives two nice properties. First, they are guaranteed to be unique; secondly they can be generated only in cases where a commit has been made successfully (within the required time period). The 
\haskellinline{do} notation supported by monads allows the generated names to be captured by the \haskellinline{bind} operation and referred to later in the computation.

\subsection{Potential model analyses}

Given that we have a language and a semantics for it, we can examine its properties. These might be properties of individual runs of a particular contract, of all runs of a given contract, or indeed properties of all the contracts of the language. In this section we look at two particularly obvious questions, but these are not the only ones that we might ask.

\paragraph{What are the valid contracts?}

One possible definition is ``those contracts that do not lead to a \haskellinline{FailedPay} action''.
Given the mechanisms for making and using commitments are limited by timeouts, it should be possible to write contracts that have this property, perhaps with some assumptions about the value of observables.
Looking at this in a little more detail, we would expect the following points to hold.
\begin{itemize}
\item For a \haskellinline{Pay} contract a sufficient balance not to produce a \haskellinline{FailedPay} or a \haskellinline{DuplicateRedeem} action is required.. 
\item For \haskellinline{Both} we require a sufficient balance for both contracts combined. Need to think here whether the two conjuncts should be independent, or whether a commitment in one half can be used in the other.
\item For \haskellinline{Choice} we require a sufficient balance for the maximum of the requirements of the individual contracts. 
\item For \haskellinline{When}, need to ``age'' the existing commitments relative to the new starting point of the continuation contract. 
\item Are there any constraints on  \haskellinline{CommitCash} and \haskellinline{RedeemCC} statements? Not directly, but there need to be sufficient funds available to pay if needed, or enough time for the commitments to time out, if not.
\end{itemize}

\paragraph{Do all contracts potentially evaluate to \haskellinline{Null}?}

That is, for each contract do there exist sequences of values of \haskellinline{Commit} and \haskellinline{OS} such that stepping the contract with these inputs leads to \haskellinline{Null}? Some relevant points include:
\begin{itemize}
\item For \haskellinline{When}, need to be able to make \haskellinline{interpretObs st obs os} true. Since there is a timeout to \haskellinline{When} then there's no need to impose this condition, as we'll progress at timeout if not before.
\item As we have put timeouts on to \haskellinline{CommitCash} and \haskellinline{RedeemCC} operations, there are no remaining requirements on these constructs for contract execution to make progress, albeit eventually.
\item We need to check the status of \haskellinline{Observations}: what if an observation is unsatisfiable?
\item There are no constraints on other constructs.
\end{itemize}

\subsection{Alternatives}

@@

\section{Response to comments on the original document made by Charles on Slack}
